Bayesian calibration of a hybrid nuclear collision model\\ using $\boldsymbol p$-Pb and Pb-Pb data from the LHC

### 2018-10-23 paper/paper.tex
Update paper revisions

\begin{abstract}
We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference. Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$, and flow cumulants.
We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a single proton.
Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton are discussed.
\end{abstract}

@@

With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free-streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$, and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.

@@

We employ a multi-stage hybrid transport model that uses relativistic viscous hydrodynamics to describe the QGP medium and microscopic Boltzmann transport to simulate the dynamics of the system after hadronization \cite{Shen:2014vra, Bernhard:2016tnd}.
The hydrodynamic initial conditions are provided by a modified version of the \trento\ model \cite{Moreland:2014oya} with additional parameters to vary the number and size of hot spots inside the nucleon.
Each initial condition profile is free-streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \cite{Broniowski:2008qk, Heinz:2015arc}.
Many of the components of the present model have been documented in previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}; we review each component here for completeness.

@@

The energy density $e$ deposited at midrapidity $\eta_s=0$ and proper time $\tau_0 \ll 1~\fmc$ is then some function,
\begin{equation}
  e(\xv_\perp, \eta_s=0, \tau_0) = f(\T_A, \T_B),
\end{equation}
of the local participant thicknesses $\T_A$ and $\T_B$.
A natural first guess for this mapping is an arithmetic mean,
\begin{equation}
  e(\xv_\perp, \eta_s=0, \tau_0) = \text{const} \times \frac{\T_A + \T_B}{2},
\end{equation}
equal to a participant or ``wounded nucleon'' model for initial energy deposition up to meaningless factor of two in the overall normalization constant.
This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for initial midrapidity energy\footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are qualitatively the same the energy density $e$ is used instead.} deposition as a function of participant density.

@@

We note that previous applications of the \trento\ model \cite{Moreland:2014oya, Bernhard:2016tnd} used the generalized mean ansatz to parametrize the \emph{entropy} density $s(\xv_\perp, \eta_s=0, \tau_\text{hydro})$, motivated (in part) by the approximate linear relation at midrapidity between initial entropy density $dS/d\eta_s$ and final charged-particle density $d\nch/d\eta_s$ \cite{Shen:2015qta}.
These studies, however, assumed instant thermalization, an approximation which sets $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ identically to zero at the hydrodynamic starting time $\tau_\text{hydro}$.
The present study, meanwhile, uses a pre-equilibrium free streaming stage to estimate $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ as a function of a variable free-streaming time $\taufs$, at which point the system is matched to hydrodynamics.
Strictly speaking, the system's entropy is ill defined prior to thermalization while the system free streams.
We therefore reinterpret the quantity produced by the \trento\ model as an energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ and free-stream this quantity to the hydrodynamic starting time $\tau_\text{hydro} \equiv \taufs$.
This assumes, of course, that the model's parametrization for entropy density is also suitable as a parametrization for the energy density.
To this point, we note that the arguments for the generalized mean ansatz are quite general \cite{Moreland:2014oya}, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta_s$.
We thus expect the parametrization to be suitable for either quantity.

@@

\begin{figure}
  \includegraphics{coupling}
  \caption{Cartoon of the free-streaming approximation for hydrodynamic initialization. The initial state is free-streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling). This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
  \label{fig:coupling}
\end{figure}

@@

There are of course two limiting cases for the strength of interactions inside the QGP medium immediately after the collision: infinitely weak coupling where the secondary partons free-stream without interacting, and infinitely strong coupling where the fluid's inter-particle mean free path effectively vanishes.
Realistically, one expects the initial parton interactions to lie somewhere between these two extremes.
We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:coupling}, which free-streams the initial state for proper time $\taufs$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \cite{Liu:2015nwa, Broniowski:2008qk}.
The free parameter $\taufs$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \taufs \lesssim1\ \fmc$.

@@

After free-streaming for proper time $\taufs$, we transition to viscous hydrodynamics which solves the conservation equations
\begin{equation}
  \label{eq:continuity}
  \partial_\mu T^{\mu\nu} = 0
\end{equation}
for the hydrodynamic energy-momentum tensor $T^{\mu\nu}$ expressed in Eq.~\eqref{hydro-eqn} using a set of second-order Israel-Stewart equations formulated in the 14-moment approximation
\cite{Israel:1979wp, Israel:1976aa, Denicol:2012cn, Denicol:2010xn}.
This produces a pair of relaxation-type equations
\begin{subequations}
  \label{eq:relaxation}
  \begin{align}
    \tau_\Pi \Pi + \dot{\Pi} &=
      - \zeta \theta - \delta_{\Pi\Pi} \Pi\theta
      + \lambda_{\Pi\pi} \pi^{\mu\nu} \sigma_{\mu\nu}, \\[1ex]
    \tau_\pi \dot{\pi}^{\langle \mu\nu \rangle} + \pi^{\mu\nu} &=
      2\eta\sigma^{\mu\nu} - \delta_{\pi\pi} \pi^{\mu\nu} \theta
      + \phi_7 \pi_\alpha^{\langle \mu} \pi^{\nu \rangle \alpha} \nonumber \\
      &\qquad {} - \tau_{\pi\pi} \pi_\alpha^{\langle \mu}\sigma^{\nu \rangle \alpha}
      + \lambda_{\pi\Pi} \Pi \sigma^{\mu\nu},
  \end{align}
\end{subequations}
for the bulk pressure $\Pi$ and shear-stress $\pi^{\mu\nu}$.
We model the shear viscosity $\eta$ and bulk viscosity $\zeta$ as unknown temperature dependent quantities and fix the remaining transport coefficients $\{\tau_\Pi, \delta_{\Pi\Pi}, \lambda_{\Pi\pi}, \tau_\pi, \delta_{\pi\pi}, \phi_7, \tau_{\pi\pi}, \lambda_{\pi\Pi}\}$ using analytic results derived in the limit of small but finite masses \cite{Denicol:2014vaa}.

@@

One primary advantage of using a microscopic transport model such as UrQMD as an afterburner, is that it realistically simulates the system break-up when the mean free path becomes large relative to the system size.
This dilute limit is expected to play a significant role in small collision systems where the produced medium is smaller and shorter lived.

@@

The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\xv$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
Given values for the parameters $\xv$, the model may be used to predict a vector of simulated observables $\yv$.
For example, $\yv$ might be a vector consisting of charged particle yields in different centrality bins.
The physics model thus represents a vector-valued function $f(\xv) = \yv$ which maps the parameter values $\xv$ to the model observables $\yv$.

@@

\begin{table}[t]
  \caption{Input parameter ranges for the physics model.}
  \begin{ruledtabular}
  \begin{tabular}{lll}
    Parameter         & Description                        & Range             \\
    \paddedhline
    Norm              & Normalization factor                 & 9--28 GeV       \\
    $p$               & Energy deposition parameter         & $-1$ to $+1$     \\
    $\sigmaf$         & Nucleon fluctuation std.\ dev.\      & 0--2            \\
    $r$               & Constituent sampling radius          & 0--1.2 fm       \\
    $n_c$             & Number of nucleon constituents       & 1--9            \\
    $v$               & Constituent width                    & 0.2--1.2 fm     \\
    $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
    $\taufs$          & Free-streaming time                  & 0.1--1.5 \fmc   \\
    $\smin$           & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
    $\sslope$         & Slope of $\eta/s$ above $T_c$        & 0--8 GeV$^{-1}$ \\
    $\scrv$           & Curvature of $\eta/s$ above $T_c$    & $-1$ to $+1$    \\
    $\bmax$           & Maximum value of $\zeta/s$           & 0--0.1          \\
    $\bwidth$         & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
    $\bloc$           & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
    $\Tsw$            & Switching/particlization temp.       & 135--165 MeV    \\
  \end{tabular}
  \end{ruledtabular}
  \label{tab:design}
\end{table}

@@

For the prior $P(H_\xv)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Table~\ref{tab:design}.
We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior.
For example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
One exception is the constituent number $n_c$ which we limit for practical considerations.
Recall that each constituent fluctuates independently, weighted by a gamma random variable.
Hence for constituent numbers $n_c \gg 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish.
To counteract this effect, the constituent fluctuation variance must increase as $n_c$ increases.
Eventually, these required fluctuations become unreasonably large.
We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence limit $n_c$ to this prior range.

The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\xv$ by comparing the model predictions $\yv(\xv)$ to experimental data $\yv_e$.
We focus on simple experimental observables in the present study which are sensitive to the bulk properties of the produced medium.
We calculate for each set of model parameters the following observables at midrapidity:
\begin{itemize}[leftmargin=1\parindent, itemsep=0pt]
  \item Charged-particle multiplicity $d\nch/d\eta$.
  \item Identified particle yields $dN/dy$ of pions, kaons, and protons.
  \item Transverse energy production $dE_T/d\eta$.
  \item Charged particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV).
  \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons, and protons.
  \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged particles, $0.15 < p_T < 2.0$~GeV).
  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged particles, $0.2 < p_T < 5.0$~GeV for ALICE, and $0.3 < p_T < 3.0$~GeV for CMS).
  \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged particles, $0.2 < p_T < 5.0$~GeV).
  \item Symmetric cumulants $\SC(4, 2)$ and $\SC(3,2)$.
\end{itemize}
Each observable is calculated from the list of final state particles produced by UrQMD using the same methods applied by experiment.
We generally match the kinematic cuts of all measurements with two exceptions: we use a larger rapidity interval $|\eta| < 0.8$ than experiment for some boost-invariant observables to improve our finite particle statistics, and we do not apply a rapidity gap, e.g.\ $|\Delta \eta| > 1$, between pairs of particles when calculating the two-particle cumulant $\vnk{n}{2}$ since we already oversample particles from each hydrodynamic event, and this oversampling suppresses non-flow correlations.

@@

Most of the calibration observables listed in Table~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
When calculating these observables, we generate \order{4} minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.

@@

For the correlation structure between different observable bins, we assert a simple Gaussian form
\begin{equation}
  \label{eq:corr}
  \rho_{ij}^\text{sys} = \exp \left[ -\frac{1}{2} \left(\frac{b_i - b_j}{l} \right)^2 \right],
\end{equation}
where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity), and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
We use centrality correlation lengths $l=100$ for all of the centrality binned Pb-Pb observables and $l=30$ for the centrality binned $p$-Pb charged particle yield $d\nch/d\eta$.
The \mbox{$p$-Pb} mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ which necessitate a smaller correlation length $l=5$.
We show an example correlation matrix
\begin{equation}
  \mathrm{corr}(y_i, y_j) = \cov(y_i, y_j)/(\sigma_i \sigma_j),
\end{equation}
for the Pb-Pb experimental data constructed using Eq.~\eqref{eq:corr} on the right-side of Fig.~\ref{fig:correlation}.
Here $y_i$ denotes an element of the experimental data $\yv_e$ and $\sigma_i$ its corresponding uncertainty.
The correlation matrix is block diagonal, with each block representing the correlations within a single class of observable.

@@

For example, using the median parameter estimates $r=0.88$~fm, $n_c=6$, and $v=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
This produces an ensemble-averaged nucleon density $\langle \rho \rangle$ of Gaussian width $w = 0.96$~fm, i.e.\

obtain an ensemble-averaged Gaussian nucleon width $w = 0.96$~fm, corresponding to a three-dimensional root-mean-square radius $R_p = \sqrt{3} w = 1.66$~fm.
This estimate is consistent with the larger nucleon width extracted by a similar Bayesian analysis lacking nucleon substructure \cite{Bernhard:2018hnz}.
This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
Those studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon, with a Gaussian width $w_g \approx 0.4$~fm which is roughly \emph{half} the Gaussian width preferred by our analysis.
Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.

>>>>

\begin{abstract}
We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference. Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged-particle yields, mean $p_T$, and flow cumulants.
We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a single proton.
Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton are discussed.
\end{abstract}

@@

With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged-particle yields, mean $p_T$, and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.

@@

We employ a multi-stage hybrid transport model that uses relativistic viscous hydrodynamics to describe the QGP medium and microscopic Boltzmann transport to simulate the dynamics of the system after hadronization \cite{Shen:2014vra, Bernhard:2016tnd}.
The hydrodynamic initial conditions are provided by a modified version of the \trento\ model \cite{Moreland:2014oya} with additional parameters to vary the number and size of hot spots inside the nucleon.
Each initial condition profile is free streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \cite{Broniowski:2008qk, Heinz:2015arc}.
Many of the components of the present model have been documented in previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}; we review each component here for completeness.

@@

The energy density $e$ deposited at midrapidity $\eta_s=0$ and proper time $\tau_0 \ll 1~\fmc$ is then some function,
\begin{equation}
  e(\xv_\perp, \eta_s=0, \tau_0) = f(\T_A, \T_B),
\end{equation}
of the local participant thicknesses $\T_A$ and $\T_B$.
A natural first guess for this mapping is an arithmetic mean,
\begin{equation}
  e(\xv_\perp, \eta_s=0, \tau_0) = \text{const} \times \frac{\T_A + \T_B}{2},
\end{equation}
equal to a participant or ``wounded nucleon'' model for initial energy deposition up to meaningless factor of two in the overall normalization constant.
This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged-particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for midrapidity energy deposition\footnote{Typically, wounded nucleon scaling is used to parametrize the entropy density $s$, but the shortcomings of the model are nevertheless the same when parametrizing the system's energy density $e$. Both parametrizations underpredict the steep rise of particle production observed in central collisions.}
as a function of participant density.

@@

We note that previous applications of the \trento\ model \cite{Moreland:2014oya, Bernhard:2016tnd} used the generalized mean ansatz to parametrize the \emph{entropy} density $s(\xv_\perp, \eta_s=0, \tau_\text{hydro})$, motivated (in part) by the approximate linear relation at midrapidity between initial entropy density $dS/d\eta_s$ and final charged-particle density $d\nch/d\eta$ \cite{Shen:2015qta}.
These studies, however, assumed instant thermalization, an approximation which sets $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ identically to zero at the hydrodynamic starting time $\tau_\text{hydro}$.
The present study, meanwhile, uses a pre-equilibrium free streaming stage to estimate $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ as a function of a variable free streaming time $\taufs$, at which point the system is matched to hydrodynamics.
Strictly speaking, the system's entropy is ill defined prior to thermalization while the system free streams.
We therefore reinterpret the quantity produced by the \trento\ model as an energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ and free stream this quantity to the hydrodynamic starting time $\tau_\text{hydro} \equiv \taufs$.
This assumes, of course, that the model's parametrization for entropy density is also suitable as a parametrization for the energy density.
To this point, we note that the arguments for the generalized mean ansatz are quite general \cite{Moreland:2014oya}, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta$.
We thus expect the parametrization to be suitable for either quantity.

@@

\begin{figure}
  \includegraphics{coupling}
  \caption{Cartoon of the free streaming approximation for hydrodynamic initialization. The initial state is free streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling). This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
  \label{fig:coupling}
\end{figure}

@@

There are of course two limiting cases for the strength of interactions inside the QGP medium immediately after the collision: infinitely weak coupling where the secondary partons free stream without interacting, and infinitely strong coupling where the fluid's inter-particle mean-free-path effectively vanishes.
Realistically, one expects the initial parton interactions to lie somewhere between these two extremes.
We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:coupling}, which free streams the initial state for proper time $\taufs$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \cite{Liu:2015nwa, Broniowski:2008qk}.
The free parameter $\taufs$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \taufs \lesssim1\ \fmc$.

@@

After free streaming for proper time $\taufs$, we transition to viscous hydrodynamics which solves the conservation equations
\begin{equation}
  \label{eq:continuity}
  \partial_\mu T^{\mu\nu} = 0
\end{equation}
for the hydrodynamic energy-momentum tensor $T^{\mu\nu}$ expressed in Eq.~\eqref{hydro-eqn} using a set of second-order Israel-Stewart equations formulated in the 14-moment approximation
\cite{Israel:1979wp, Israel:1976aa, Denicol:2012cn, Denicol:2010xn}.
This produces a pair of relaxation-type equations
\begin{subequations}
  \label{eq:relaxation}
  \begin{align}
    \tau_\Pi \Pi + \dot{\Pi} &=
      - \zeta \theta - \delta_{\Pi\Pi} \Pi\theta
      + \lambda_{\Pi\pi} \pi^{\mu\nu} \sigma_{\mu\nu}, \\[1ex]
    \tau_\pi \dot{\pi}^{\langle \mu\nu \rangle} + \pi^{\mu\nu} &=
      2\eta\sigma^{\mu\nu} - \delta_{\pi\pi} \pi^{\mu\nu} \theta
      + \phi_7 \pi_\alpha^{\langle \mu} \pi^{\nu \rangle \alpha} \nonumber \\
      &\qquad {} - \tau_{\pi\pi} \pi_\alpha^{\langle \mu}\sigma^{\nu \rangle \alpha}
      + \lambda_{\pi\Pi} \Pi \sigma^{\mu\nu},
  \end{align}
\end{subequations}
for the bulk pressure $\Pi$ and shear-stress $\pi^{\mu\nu}$.
We model the shear viscosity $\eta$ and bulk viscosity $\zeta$ as unknown temperature dependent quantities and fix the remaining transport coefficients $\{\tau_\Pi, \delta_{\Pi\Pi}, \lambda_{\Pi\pi}, \tau_\pi, \delta_{\pi\pi}, \phi_7, \tau_{\pi\pi}, \lambda_{\pi\Pi}\}$ using analytic results derived in the limit of small but finite masses \cite{Denicol:2014vaa}.

@@

One primary advantage of using a microscopic transport model such as UrQMD as an afterburner, is that it realistically simulates the system break-up when the mean-free-path becomes large relative to the system size.
This dilute limit is expected to play a significant role in small collision systems where the produced medium is smaller and shorter lived.

@@

The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\xv$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
Given values for the parameters $\xv$, the model may be used to predict a vector of simulated observables $\yv$.
For example, $\yv$ might be a vector consisting of charged-particle yields in different centrality bins.
The physics model thus represents a vector-valued function $f(\xv) = \yv$ which maps the parameter values $\xv$ to the model observables $\yv$.

@@

\begin{table}[t]
  \caption{Input parameter ranges for the physics model.}
  \begin{ruledtabular}
  \begin{tabular}{lll}
    Parameter         & Description                        & Range             \\
    \paddedhline
    Norm              & Normalization factor                 & 9--28 GeV       \\
    $p$               & Energy deposition parameter         & $-1$ to $+1$     \\
    $\sigmaf$         & Nucleon fluctuation std.\ dev.\      & 0--2            \\
    $r$               & Constituent sampling radius          & 0--1.2 fm       \\
    $n_c$             & Number of nucleon constituents       & 1--9            \\
    $v$               & Constituent width                    & 0.2--1.2 fm     \\
    $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
    $\taufs$          & Free streaming time                  & 0.1--1.5 \fmc   \\
    $\smin$           & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
    $\sslope$         & Slope of $\eta/s$ above $T_c$        & 0--8 GeV$^{-1}$ \\
    $\scrv$           & Curvature of $\eta/s$ above $T_c$    & $-1$ to $+1$    \\
    $\bmax$           & Maximum value of $\zeta/s$           & 0--0.1          \\
    $\bwidth$         & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
    $\bloc$           & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
    $\Tsw$            & Switching/particlization temp.       & 135--165 MeV    \\
  \end{tabular}
  \end{ruledtabular}
  \label{tab:design}
\end{table}

@@

For the prior $P(H_\xv)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Table~\ref{tab:design}.
We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior.
For example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
One exception is the constituent number $n_c$ which we limit for practical considerations.
Recall that each constituent fluctuates independently, weighted by a gamma random variable.
Hence for constituent numbers $n_c \gg 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish.
To counteract this effect, the constituent fluctuation variance must increase as $n_c$ increases.
Eventually, these required fluctuations become unreasonably large.
We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence we limit $n_c$ to this prior range.

The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\xv$ by comparing the model predictions $\yv(\xv)$ to experimental data $\yv_e$.
We focus on simple experimental observables in the present study which are sensitive to the bulk properties of the produced medium.
We calculate for each set of model parameters the following observables at midrapidity:
\begin{itemize}[leftmargin=1\parindent, itemsep=0pt]
  \item Charged-particle multiplicity $d\nch/d\eta$.
  \item Identified particle yields $dN/dy$ of pions, kaons, and protons.
  \item Transverse energy production $dE_T/d\eta$.
  \item Charged-particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV).
  \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons, and protons.
  \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged-particles, $0.15 < p_T < 2.0$~GeV).
  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged-particles, $0.2 < p_T < 5.0$~GeV for ALICE, and $0.3 < p_T < 3.0$~GeV for CMS).
  \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged-particles, $0.2 < p_T < 5.0$~GeV).
  \item Symmetric cumulants $\SC(4, 2)$ and $\SC(3,2)$.
\end{itemize}
Each observable is calculated from the list of final state particles produced by UrQMD using the same methods applied by experiment.
We generally match the kinematic cuts of all measurements with two exceptions: we use a larger rapidity interval $|\eta| < 0.8$ than experiment for some boost-invariant observables to improve our finite particle statistics, and we do not apply a rapidity gap, e.g.\ $|\Delta \eta| > 1$, between pairs of particles when calculating the two-particle cumulant $\vnk{n}{2}$ since we already oversample particles from each hydrodynamic event, and this oversampling suppresses non-flow correlations.

@@

Most of the calibration observables listed in Table~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged-particle yield in a given rapidity window.
When calculating these observables, we generate \order{4} minimum-bias events at each design point and divide the events into centrality bins using the charged-particle yield at midrapidity, similar to the procedure used by experiment.

@@

For the correlation structure between different observable bins, we assert a simple Gaussian form
\begin{equation}
  \label{eq:corr}
  \rho_{ij}^\text{sys} = \exp \left[ -\frac{1}{2} \left(\frac{b_i - b_j}{l} \right)^2 \right],
\end{equation}
where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity), and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
We use centrality correlation lengths $l=100$ for all of the centrality binned Pb-Pb observables and $l=30$ for the centrality binned $p$-Pb charged-particle yield $d\nch/d\eta$.
The \mbox{$p$-Pb} mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ which necessitate a smaller correlation length $l=5$.
We show an example correlation matrix
\begin{equation}
  \mathrm{corr}(y_i, y_j) = \cov(y_i, y_j)/(\sigma_i \sigma_j),
\end{equation}
for the Pb-Pb experimental data constructed using Eq.~\eqref{eq:corr} on the right-side of Fig.~\ref{fig:correlation}.
Here $y_i$ denotes an element of the experimental data $\yv_e$ and $\sigma_i$ its corresponding uncertainty.
The correlation matrix is block diagonal, with each block representing the correlations within a single class of observable.

@@

For example, using the posterior distribution's median values, $r=0.88$~fm, $n_c=6$, and $v=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
The resulting ensemble-averaged nucleon density,
\begin{equation}
  \bar{\rho}_n(\xv) = \frac{1}{(2 \pi w^2)^{3/2}} \exp \left( -\frac{|\xv|^2}{2 w^2} \right),
\end{equation}
is described by a single Gaussian of width $w = 0.96$~fm.
This nucleon width is consistent with a previous estimate, $w = 0.96_{-0.05}^{+0.04}$~fm, obtained by a similar Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV using a physics model without nucleon substructure \cite{Bernhard:2018hnz}.

This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
Those studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon, with a Gaussian width $w_g \approx 0.4$~fm which is roughly \emph{half} the Gaussian width preferred by our analysis.
Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.

Moving on, we direct our attention to the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}.
The distribution is not sharply peaked, and hence we refrain from quoting a distribution median and 90\% credible interval.
Note, however, that the posterior clearly favors $n_c > 1$ constituents.
This is not surprising.
The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
